### 2021-05-07 manuscript.tex
added conclusion

\section{Abstract}
Maintenance and upgrades are a fundamental requirement of any software-based tool, and the resources within the field of Bioinformatics are no exception. Too often, the combination of self-taught programmers in former Biologists and the broader lack of a software development ecosystem can propagate elevated degrees of software entropy, also known as technical debt. It is then a fair assumption to say that this level of entropy is higher than in fields which more consistently employ the skills of professional developers. Here, we present the case of the Candidate Cancer Gene Database as a minor instance of taking advantage of an upgrade cycle in an attempt to service some of this technical debt. The goals of this project ultimately strived to acheive the implementation of a new, core framework at the base of the data integration backend as well as a more modern and simplified web front-end. In all, this represented a complete rebuild of the application while preserving the existing conceptual implementation. Our work served to establish a new application with greater structural simplicity, improved resource use, and, most relevant, a code base which will be easier to maintain and upgrade in the future.

@@

\begin{figure}[H]
    \center{}
    \includegraphics[width=0.8\textwidth]{fig/file_tree.pdf}
    \caption[Project Directory File Tree Comparison]{\textbf{Project Directory File Tree Comparison} The above figure offers a qualitative assessment of an element of technical debt that was improved in refactoring. Shown are the file trees of the site directories of the new and old projects. What is illustrated is a dramatic improvement in the tree size and complexity through the elimination of a large degree of content, log, and style files. This is accentuated even further when comparing the source code. This was accomplished both in the intentional design of eliminating content redundancy and in the streamlined structure of the Rmarkdown rendering process. This allowed for the elimination of several dimensions of broader web design and table build data elements.}\label{fig:fileTree}
\end{figure}

@@

\chapter{Conclusions}

Best practices. Discuss open source.

Mention cosmic connection.

>>>>

\section{Abstract}
Maintenance and upgrades are a fundamental requirement of any software-based tool, and the resources within the field of Bioinformatics are no exception. Too often, the combination of using self-taught programmers who are typically trained Biologists along with the broader lack of a software development ecosystem can propagate elevated degrees of software entropy, also known as technical debt. It is then a fair assumption to say that this level of entropy is higher than in fields which more consistently employ the skills of professional developers. Here, we present the case of the Candidate Cancer Gene Database (CCGD, https://z.umn.edu/ccgd) as a minor instance of taking advantage of an upgrade cycle in an attempt to service some of this technical debt. The goals of this project ultimately strived to acheive the implementation of a new, core framework at the base of the data integration backend as well as a more modern and simplified web front-end. In all, this represented a complete rebuild of the application while preserving the existing conceptual implementation. Our work served to establish a new application with greater structural simplicity, improved resource use, and, most relevant, a code base which will be easier to maintain and upgrade in the future.

@@

\begin{figure}[H]
    \center{}
    \includegraphics[width=0.8\textwidth]{fig/file_tree.pdf}
    \caption[Project Directory File Tree Comparison]{\textbf{Project Directory File Tree Comparison} The above figure offers a qualitative assessment of an element of technical debt that was improved in refactoring. Shown are the horizontally oriented file trees of the site directories of the new and old projects. What is illustrated is a dramatic improvement in the tree size and complexity through the elimination of a large degree of redudant or unecessary content, log, and style files. This is accentuated even further when comparing the source code. This was accomplished both in the intentional design of eliminating content redundancy and in the streamlined structure of the Rmarkdown rendering process. This allowed for the elimination of several dimensions of broader web design and table build data elements.}\label{fig:fileTree}
\end{figure}

@@

\chapter{Conclusions}
This work has given new life to the CCGD and the valuable functionality that it offers in the form of accessible forward genetic cancer screen data. Further, it has been a case study in the benefits of evolving of tool sets and frameworks that increasingly live at the nexus of the field of Bioinformatics. We've given a general characterization of the contrast that can exist between development schemas that are separated by only 5 years time and the substantive impact to be had in adopting these new paradigms.

With that, the CCGD will continue to function as it has since its initial release by offering a point of integration for collated results in the cancer gene screening space. Existing connections have been maintained both to and from external resources such as those from the CCGD pointing to the GeneCards suite, NCBI's HomoloGene, and any associated screening publication. Additionally, an external gene-level reference link pointing to the CCGD from the Sanger COSMIC database has been maintained as well. Finally, new studies will continue to be integrated into the core table as they are published and curated, as was done before.

Finally, we have maintained the essence of that functionality while notably reducing the digital entropy within the codebase and application architecture. At the heart of this redesign, we have placed the Rmarkdown web paradigm at the center of the application. We feel this offers not only an expanded repertoire of functionality and features but brings the application stack to parity with many of the comparable Bioinformatics web tools of the day. In doing so, we implemented our design with an elevated consideration for best practices in development resulting in improved resource outcomes and overall performance. This is reflected by qualitative and quantitative comparisons of the codebase. We also placed added focus on the nature of the documentation and portability of the code, which is now publicly hosted under the open-source MIT license at github (https://github.com/ctastad/ccgd), (https://doi.org/10.5281/zenodo.4422026). With this approach, we hope that the sustained service of the CCGD will continue to serve the research community at-large now and for future refactorizations yet to come.
